# By Scenario {-}

```{r Scenarios}
scenarios <- c("h_cpol", "h_ndc", "o_1p5c", "o_2c", "d_delfrag", "d_rap")
```

# Socio-Economics

## GDP

#### GDP|MER {-}

```{r}
items <- c("GDP|MER")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### GDP|PPP {-}

```{r}
items <- c("GDP|PPP")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Policy Cost

### Policy Cost|GDP Loss {-}

```{r}
items <- c("Policy Cost|GDP Loss")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Policy Cost|Consumption Loss {-}

```{r}
items <- c("Policy Cost|Consumption Loss")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Population

```{r}
items <- c("Population")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Carbon Price

```{r}
items <- c("Price|Carbon")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s, scales="fixed")
}
```

# Climate and Emissions

## Global Mean Temperature Change

```{r}
items <- c("Temperature|Global Mean")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Greenhouse Gas Emissions

### Emissions|Kyoto Gases {-}

```{r}
items <- c("Emissions|Kyoto Gases", "Emi|GHG")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Emissions|CO2 {-}

```{r}
items <- c("Emissions|CO2", "Emi|CO2")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Emissions|CO2|Energy and Industrial Processes {-}

```{r}
items <- c("Emissions|CO2|Energy and Industrial Processes", "Emi|CO2|Energy and Industrial Processes")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Emissions|CO2|AFOLU {-}

```{r}
items <- c("Emissions|CO2|AFOLU", "Emi|CO2|Land-Use Change")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Emissions|CH4 {-}

```{r}
items <- c("Emissions|CH4", "Emi|CH4")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Emissions|N2O {-}

```{r}
items <- c("Emissions|N2O", "Emi|N2O")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Emissions|F-Gases {-}

```{r}
items <- c("Emissions|F-Gases", "Emi|F-Gases")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Carbon Sequestration

### Carbon Sequestration|CCS {-}

```{r}
items <- c("Carbon Sequestration|CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Carbon Sequestration|CCS|Fossil {-}

```{r}
items <- c("Carbon Sequestration|CCS|Fossil")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Carbon Sequestration|CCS|Biomass {-}

```{r}
items <- c("Carbon Sequestration|CCS|Biomass")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Carbon Sequestration|Land Use {-}

```{r}
items <- c("Carbon Sequestration|Land Use")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```


# Energy System

## Primary Energy

```{r}
items <- c("Primary Energy", "PE")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy|Oil {-}

```{r}
items <- c("Primary Energy|Oil", "PE|Oil")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy|Coal {-}

```{r}
items <- c("Primary Energy|Coal", "PE|Coal")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy|Gas {-}

```{r}
items <- c("Primary Energy|Gas", "PE|Gas")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy|Nuclear {-}

```{r}
items <- c("Primary Energy|Nuclear", "PE|Nuclear")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy|Biomass {-}

```{r}
items <- c("Primary Energy|Biomass", "PE|Biomass")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy|Non-Biomass Renewables {-}

```{r}
items <- c("Primary Energy|Non-Biomass Renewables", "PE|Non-Biomass Renewables")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Non-Biomass Renewables|Solar {-}

```{r}
items <- c("Primary Energy|Non-Biomass Renewables|Solar", "PE|Solar")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Non-Biomass Renewables|Wind {-}

```{r}
items <- c("Primary Energy|Non-Biomass Renewables|Wind", "PE|Wind")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```


## Secondary Energy
```{r}
items <- c("Secondary Energy")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Secondary Energy|Electricity
```{r}
items <- c("Secondary Energy|Electricity", "SE|Electricity")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Coal|w/ CCS {-}
```{r}
items <- c("Secondary Energy|Electricity|Coal|w/ CCS", "SE|Electricity|Coal|w/ CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Coal|w/o CCS {-}
```{r}
items <- c("Secondary Energy|Electricity|Coal|w/o CCS", "SE|Electricity|Coal|w/o CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Gas|w/ CCS {-}
```{r}
items <- c("Secondary Energy|Electricity|Gas|w/ CCS", "SE|Electricity|Gas|w/ CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Gas|w/o CCS {-}
```{r}
items <- c("Secondary Energy|Electricity|Gas|w/o CCS", "SE|Electricity|Gas|w/o CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Oil|w/o CCS {-}
```{r}
items <- c("Secondary Energy|Electricity|Oil|w/o CCS", "SE|Electricity|Oil|w/o CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Biomass|w/ CCS {-}
```{r}
items <- c("Secondary Energy|Electricity|Biomass|w/ CCS", "SE|Electricity|Biomass|w/ CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity| {-}
```{r}
items <- c("Secondary Energy|Electricity|Biomass|w/o CCS", "SE|Electricity|Biomass|w/o CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Nuclear {-}
```{r}
items <- c("Secondary Energy|Electricity|Nuclear", "SE|Electricity|Nuclear")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Solar {-}
```{r}
items <- c("Secondary Energy|Electricity|Solar", "SE|Electricity|Solar")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Secondary Energy|Electricity|Wind {-}
```{r}
items <- c("Secondary Energy|Electricity|Wind", "SE|Electricity|Wind")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```


## Capacity Additions

### Capacity Additions|Electricity|Coal {-}

```{r}
items <- c("Capacity Additions|Electricity|Coal")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Capacity Additions|Electricity|Solar|PV {-}

```{r}
items <- c("Capacity Additions|Electricity|Solar|PV")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Final Energy

```{r}
items <- c("Final Energy", "FE")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Industry {-}

```{r}
items <- c("Final Energy|Industry", "FE|Industry")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Transportation {-}

```{r}
items <- c("Final Energy|Transportation", "FE|Transport")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Residential and Commercial {-}

```{r}
items <- c("Final Energy|Residential and Commercial", "FE|Buildings")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Electricity {-}

```{r}
items <- c("Final Energy|Electricity", "FE|Electricity")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Gases {-}

```{r}
items <- c("Final Energy|Gases", "FE|Gases")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Hydrogen {-}

```{r}
items <- c("Final Energy|Hydrogen", "FE|Hydrogen")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Liquids {-}

```{r}
items <- c("Final Energy|Liquids", "FE|Liquids")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Final Energy|Solids {-}

```{r}
items <- c("Final Energy|Solids", "FE|Solids")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```
