# IEA Net-Zero Comparison {-}

# Socio-Economics

## GDP
```{r}
plot_default(ivar="GDP|MER", ititle="GDP|MER in 2030 [billion US$2010/yr]", fname="GDP_MER_2030", iper=2030,irange=c(80000, 120000))
plot_default(ivar="GDP|MER", ititle="GDP|MER in 2050 [billion US$2010/yr]", fname="GDP_MER_2050")
plot_default(ivar="GDP|PPP", ititle="GDP|PPP in 2030 [billion US$2010/yr]", fname="GDP_PPP_2030", iper=2030, irange=c(120000, 180000))
plot_default(ivar="GDP|PPP", ititle="GDP|PPP in 2050 [billion US$2010/yr]", fname="GDP_PPP_2050", irange=c(200000, 300000))
```

## Population
```{r}
plot_default(ivar="Population", ititle="Population in 2030 [million]", fname="pop_2030", irange=c(8000, 8500), iper=2030)
plot_default(ivar="Population", ititle="Population in 2050 [million]", fname="pop_2050", irange=c(9000, 9500))
```


## Carbon Price
```{r}
plot_default(ivar="Price|Carbon", ititle="Carbon Price in 2030 [US$2010/t CO2]", fname="carbon_price_2030", irange=c(0, 1000), iper=2030)
plot_default(ivar="Price|Carbon", ititle="Carbon Price in 2050 [US$2010/t CO2]", fname="carbon_price_2050", irange=c(0, 2000))
```


# Climate and Emissions

## Greenhouse Gas Emissions

### Kyoto Gases {-}

```{r}
plot_default(ivar="Emissions|Kyoto Gases", ititle="Emissions of Kyoto Gases in 2050 [GtCO2eq/yr]", 
             fname="emi_kyoto_2030", irange = c(15, 45), iper=2030, ifac=0.001)
plot_default(ivar="Emissions|Kyoto Gases", ititle="Emissions of Kyoto Gases in 2050 [GtCO2eq/yr]", 
             fname="emi_kyoto_2050", irange = c(0, 18), ifac=0.001)
```

### CO2 {-}

```{r}
plot_default(ivar="Emissions|CO2", ititle="CO2 Emissions in 2030 [GtCO2/yr]", 
             fname="emi_co2_2030", irange = c(10, 30), iper=2030, ifac=0.001)
plot_default(ivar="Emissions|CO2", ititle="CO2 Emissions in 2050 [GtCO2/yr]", 
             fname="emi_co2_2050", irange = c(-3, 9), ifac=0.001)
```

#### Energy {-}

```{r}
plot_default(ivar="Emissions|CO2|Energy", ititle="CO2 Energy Emissions in 2030 [GtCO2/yr]", 
             fname="emi_co2_energy_2030", iper=2030, ifac=0.001)
plot_default(ivar="Emissions|CO2|Energy", ititle="CO2 Energy Emissions in 2050 [GtCO2/yr]", 
             fname="emi_co2_energy_2050", ifac=0.001)
```

#### Electricity {-}

```{r}
plot_default(ivar="Emissions|CO2|Energy|Supply|Electricity", ititle="CO2 from Electricity Production in 2030 [GtCO2/yr]", 
             fname="emi_co2_elec_2030", iper=2030, ifac=0.001)
plot_default(ivar="Emissions|CO2|Energy|Supply|Electricity", ititle="CO2 from Electricity Production in 2050 [GtCO2/yr]", 
             fname="emi_co2_elec_2050", ifac=0.001)
```

#### Industry {-}

```{r}
plot_default(ivar="Emissions|CO2|Energy|Demand|Industry", ititle="CO2 Emissions from Industry in 2030 [GtCO2/yr]", 
             fname="emi_co2_industry_2030", iper=2030, ifac=0.001)
plot_default(ivar="Emissions|CO2|Energy|Demand|Industry", ititle="CO2 Emissions from Industry in 2050 [GtCO2/yr]", 
             fname="emi_co2_industry_2050", ifac=0.001)
```

#### Transportation {-}

```{r}
plot_default(ivar="Emissions|CO2|Energy|Demand|Transportation", ititle="CO2 Emissions from Transport in 2030 [GtCO2/yr]", 
             fname="emi_co2_transp_2030", iper=2030, ifac=0.001)
plot_default(ivar="Emissions|CO2|Energy|Demand|Transportation", ititle="CO2 Emissions from Transport in 2050 [GtCO2/yr]", 
             fname="emi_co2_transp_2050", ifac=0.001)
```

#### Residential and Commercial {-}

```{r}
plot_default(ivar="Emissions|CO2|Energy|Demand|Residential and Commercial", ititle="CO2 Emissions from Buildings in 2030 [GtCO2/yr]", 
             fname="emi_co2_build_2030", iper=2030, ifac=0.001)
plot_default(ivar="Emissions|CO2|Energy|Demand|Residential and Commercial", ititle="CO2 Emissions from Buildings in 2050 [GtCO2/yr]", 
             fname="emi_co2_build_2050", ifac=0.001)
```


## Carbon Sequestration
```{r}
plot_default(ivar="Carbon Sequestration|CCS", ititle="CCS in 2030 [GtCO2/yr]", 
             fname="CCS_2030", ifac=1e-3, iper=2030, irange = c(0, 6))
plot_default(ivar="Carbon Sequestration|CCS", ititle="CCS in 2050 [GtCO2/yr]", fname="CCS_2050", ifac=1e-3)
```

### Afforestation & BECCS

```{r}
# Use of carbon dioxide removal technologies: CO2 emissions captured and stored from bioenergy with
# CCUS (BECCS) and direct air capture in the IPCC scenarios range from 3.5 Gt CO2 to 16 Gt CO2 in 2050,
# compared with 1.9 Gt CO2 in the NZE Scenario.
tmp_sr15 <- data_sr15 %>%
  filter(period == 2030,
         variable %in% c("Carbon Sequestration|CCS|Biomass", "Carbon Sequestration|Land Use|Afforestation"),
         category %in% c("Below 1.5°C", "1.5C low overshoot")) %>%
  mutate(value = value*1e-3) %>%
  group_by(model, scenario, region, period) %>%
  summarise(value = sum(value, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(source = "1 IPCC SR1.5")
tmp_iea <- data.frame(model="IEA WEO 2021 Net2050", scenario = "NZE2050", source="2 IEA", period=2030, value=1.9, region="World", stringsAsFactors = FALSE)
tmp_ngfs <- data_ngfs_world %>%
  filter(period == 2030,
         scenario %in% u_ngfs_scenario,
         variable %in% c("Carbon Sequestration|CCS|Biomass", "Carbon Sequestration|Land Use|Afforestation")) %>%
  mutate(value = value*1e-3) %>%
  select(model, scenario, region, variable, period, value) %>%
  group_by(model, scenario, region, period) %>%
  summarise(value = sum(value, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(source = "3 NGFS")

tmp_ar6 <- data_ar6 %>%
  filter(period == 2030,
         variable %in% c("Carbon Sequestration|CCS|Biomass", "Carbon Sequestration|Land Use|Afforestation"),
         category %in% c("C1")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(source = "2 IPCC AR6") %>%
  mutate(value = (`Carbon Sequestration|CCS|Biomass` + `Carbon Sequestration|Land Use|Afforestation`)*1e-3)
tmp_ar6 <- tmp_ar6 %>% filter(!is.na(value))

tmp_title <- "Carbon dioxide removal (Afforestation & BECCS) \nin 2030 [GtCO2/yr]"

p7 <- plot_comparison_sr15_ngfs_iea(tmp_sr15, tmp_ngfs, tmp_iea, tmp_ar6, tmp_title)
plot(p7)
ggsave(paste0("plot_CDR_Aff_BECCS_",tmstmp,".png"), p7 + theme(legend.position = "none"), width = 10, height = 10)


tmp_sr15 <- data_sr15 %>%
  filter(period == 2050,
         variable %in% c("Carbon Sequestration|CCS|Biomass", "Carbon Sequestration|Land Use|Afforestation"),
         category %in% c("Below 1.5°C", "1.5C low overshoot")) %>%
  mutate(value = value*1e-3) %>%
  group_by(model, scenario, region, period) %>%
  summarise(value = sum(value, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(source = "1 IPCC SR1.5")
tmp_iea <- data.frame(model="IEA WEO 2021 Net2050", scenario = "NZE2050", source="2 IEA", period=2050, value=1.9, region="World", stringsAsFactors = FALSE)
tmp_ngfs <- data_ngfs_world %>%
  filter(period == 2050,
         scenario %in% u_ngfs_scenario,
         variable %in% c("Carbon Sequestration|CCS|Biomass", "Carbon Sequestration|Land Use|Afforestation")) %>%
  mutate(value = value*1e-3) %>%
  select(model, scenario, region, variable, period, value) %>%
  group_by(model, scenario, region, period) %>%
  summarise(value = sum(value, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(source = "3 NGFS")

tmp_ar6 <- data_ar6 %>%
  filter(period == 2050,
         variable %in% c("Carbon Sequestration|CCS|Biomass", "Carbon Sequestration|Land Use|Afforestation"),
         category %in% c("C1")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(source = "2 IPCC AR6") %>%
  mutate(value = (`Carbon Sequestration|CCS|Biomass` + `Carbon Sequestration|Land Use|Afforestation`)*1e-3)
tmp_ar6 <- tmp_ar6 %>% filter(!is.na(value))

tmp_title <- "Carbon dioxide removal (Afforestation & BECCS) \nin 2050 [GtCO2/yr]"

p7 <- plot_comparison_sr15_ngfs_iea(tmp_sr15, tmp_ngfs, tmp_iea, tmp_ar6, tmp_title)
plot(p7)
ggsave(paste0("plot_CDR_Aff_BECCS_",tmstmp,".png"), p7 + theme(legend.position = "none"), width = 10, height = 10)
```

# Energy

## Primary Energy

### Fossil {-}
```{r}
plot_default(ivar="Primary Energy|Fossil", ititle="Total Fossils Use in 2030 [EJ/yr]", 
             fname="PE_fossil_2030", iper = 2030)
plot_default(ivar="Primary Energy|Fossil", ititle="Total Fossils Use in 2050 [EJ/yr]", 
             fname="PE_fossil_2050")
```

### Coal {-}
```{r}
plot_default(ivar="Primary Energy|Coal", ititle="Coal Use in 2030 [EJ/yr]", fname="PE_coal_2030", iper=2030)
plot_default(ivar="Primary Energy|Coal", ititle="Coal Use in 2050 [EJ/yr]", fname="PE_coal_2050")
```

### Oil {-}
```{r}
plot_default(ivar="Primary Energy|Oil", ititle="Oil Use in 2030 [EJ/yr]", fname="PE_oil_2030", iper=2030)
plot_default(ivar="Primary Energy|Oil", ititle="Oil Use in 2050 [EJ/yr]", fname="PE_oil_2050")
```

### Gas {-}
```{r}
plot_default(ivar="Primary Energy|Gas", ititle="Gas Use in 2030 [EJ/yr]", fname="PE_gas_2030", iper=2030)
plot_default(ivar="Primary Energy|Gas", ititle="Gas Use in 2050 [EJ/yr]", fname="PE_gas_2050")
```

### Biomass {-}
```{r}
plot_default(ivar="Primary Energy|Biomass", ititle="Primary Bioenergy in 2030 [EJ/yr]", 
             fname="PE_biomass_2030", irange=c(50, 150), iper=2030)
plot_default(ivar="Primary Energy|Biomass", ititle="Primary Bioenergy in 2050 [EJ/yr]", 
             fname="PE_biomass_2050", irange=c(50, 250))
```

### Non-Biomass Renewables {-}
```{r}
plot_default(ivar="Primary Energy|Non-Biomass Renewables", ititle="Primary Energy Non-Biomass Renewables in 2030 [EJ/yr]", 
             fname="PE_non_biomass_2030", irange=c(0, 150), iper=2030)
plot_default(ivar="Primary Energy|Non-Biomass Renewables", ititle="Primary Energy Non-Biomass Renewables in 2050 [EJ/yr]", 
             fname="PE_non_biomass_2050", irange=c(0, 400))
```

### Nuclear {-}
```{r}
plot_default(ivar="Primary Energy|Nuclear", ititle="Primary Energy Nuclear in 2030 [EJ/yr]", 
             fname="PE_nuclear_2030", irange=c(0, 35), iper=2030)
plot_default(ivar="Primary Energy|Nuclear", ititle="Primary Energy Nuclear in 2050 [EJ/yr]", 
             fname="PE_nuclear_2050", irange=c(0, 60))
```

## Final Energy
```{r}
plot_default(ivar="Final Energy",ititle="Final Energy Use in 2030 [EJ/yr]", fname="FE_2030", irange=c(250,500), iper=2030)
plot_default(ivar="Final Energy",ititle="Final Energy Use in 2050 [EJ/yr]", fname="FE_2050", irange=c(250,500))
```

### Hydrogen {-}
```{r}
plot_default(ivar="Final Energy|Hydrogen",ititle="Hydrogen Use in 2030 [EJ/yr]", fname="FE_hydrogen_2030", iper=2030)
plot_default(ivar="Final Energy|Hydrogen",ititle="Hydrogen Use in 2050 [EJ/yr]", fname="FE_hydrogen_2050")
```

## Electricity Shares Solar/Wind
```{r}
tmp_sr15 <- data_sr15 %>%
  filter(period == 2030,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar"),
         category %in% c("Below 1.5°C", "1.5C low overshoot")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)# IEA data taken from NZE2021_Annex.xlsb > World_Elec > (Solar PV + Wind ü CSP) / Total generation  * 100
tmp_ar6 <- data_ar6 %>%
  filter(period == 2030,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar"),
         category %in% c("C1")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)# IEA data taken from NZE2021_Annex.xlsb > World_Elec > (Solar PV + Wind ü CSP) / Total generation  * 100
tmp_ar6 <- tmp_ar6 %>% filter(!is.na(value))

tmp_iea <- data_iea_nz %>%
  filter(period == 2030,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)

tmp_ngfs <- data_ngfs_world %>%
  filter(period == 2030,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar"),
         scenario %in% u_ngfs_scenario) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)
tmp_title <- "Shares of wind and solar in electricity generation in 2030 [%]"

p1 <- plot_comparison_sr15_ngfs_iea(tmp_sr15, tmp_ngfs, tmp_iea, tmp_ar6, tmp_title)
print(p1)
ggsave(paste0("0_plot_Elec_Share_Solar_Wind_",tmstmp,"_2030_withLegend.png"), p1, width = 10, height = 10)



tmp_sr15 <- data_sr15 %>%
  filter(period == 2050,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar"),
         category %in% c("Below 1.5°C", "1.5C low overshoot")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)# IEA data taken from NZE2021_Annex.xlsb > World_Elec > (Solar PV + Wind ü CSP) / Total generation  * 100
tmp_ar6 <- data_ar6 %>%
  filter(period == 2050,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar"),
         category %in% c("C1")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)# IEA data taken from NZE2021_Annex.xlsb > World_Elec > (Solar PV + Wind ü CSP) / Total generation  * 100
tmp_ar6 <- tmp_ar6 %>% filter(!is.na(value))

tmp_iea <- data_iea_nz %>%
  filter(period == 2050,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar")) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)

tmp_ngfs <- data_ngfs_world %>%
  filter(period == 2050,
         variable %in% c("Secondary Energy|Electricity", "Secondary Energy|Electricity|Wind", "Secondary Energy|Electricity|Solar"),
         scenario %in% u_ngfs_scenario) %>%
  select(model, scenario, region, variable, period, value) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  mutate(variable = "", unit="") %>%
  mutate(value = 100 * (`Secondary Energy|Electricity|Wind` + `Secondary Energy|Electricity|Solar`) / `Secondary Energy|Electricity`)
tmp_title <- "Shares of wind and solar in electricity generation in 2050 [%]"

p1 <- plot_comparison_sr15_ngfs_iea(tmp_sr15, tmp_ngfs, tmp_iea, tmp_ar6, tmp_title)
print(p1)
ggsave(paste0("0_plot_Elec_Share_Solar_Wind_",tmstmp,"_withLegend.png"), p1, width = 10, height = 10)

```
