# By Scenario
```{r Scenarios}
scenarios <- c("d_delfrag", "d_rap", "h_cpol", "h_ndc", "o_1p5c", "o_2c")
```

## Socio-Economics

### GDP

#### GDP|MER

```{r}
items <- c("GDP|MER")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### GDP|PPP

```{r}
items <- c("GDP|PPP")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Policy Cost

#### Policy Cost|GDP Loss

```{r}
items <- c("Policy Cost|GDP Loss")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Policy Cost|Consumption Loss

```{r}
items <- c("Policy Cost|Consumption Loss")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Population

```{r}
items <- c("Population")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Carbon Price

```{r}
items <- c("Price|Carbon")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

## Climate and Emissions

### Global Mean Temperature Change

```{r}
items <- c("Temperature|Global Mean")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Greenhouse Gas Emissions

#### Emissions|Kyoto Gases

```{r}
items <- c("Emissions|Kyoto Gases", "Emi|GHG")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### CO2 Emissions

```{r}
items <- c("Emissions|CO2", "Emi|CO2")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

##### Emissions|CO2|Energy and Industrial Processes

```{r}
items <- c("Emissions|CO2|Energy and Industrial Processes", "Emi|CO2|Energy and Industrial Processes")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

##### Emissions|CO2|AFOLU

```{r}
items <- c("Emissions|CO2|AFOLU", "Emi|CO2|Land-Use Change")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Emissions|CH4

```{r}
items <- c("Emissions|CH4", "Emi|CH4")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Emissions|N2O

```{r}
items <- c("Emissions|N2O", "Emi|N2O")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Emissions|F-Gases

```{r}
items <- c("Emissions|F-Gases", "Emi|F-Gases")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Carbon Sequestration

#### Carbon Sequestration|CCS

```{r}
items <- c("Carbon Sequestration|CCS")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

##### Carbon Sequestration|CCS|Fossil

```{r}
items <- c("Carbon Sequestration|CCS|Fossil")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

##### Carbon Sequestration|CCS|Biomass

```{r}
items <- c("Carbon Sequestration|CCS|Biomass")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Carbon Sequestration|Land Use

```{r}
items <- c("Carbon Sequestration|Land Use")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```


## Energy System

### Final Energy

```{r}
items <- c("Final Energy", "FE")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Final Energy|Industry

```{r}
items <- c("Final Energy|Industry", "FE|Industry")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Final Energy|Transportation

```{r}
items <- c("Final Energy|Transportation", "FE|Transport")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Final Energy|Residential and Commercial

```{r}
items <- c("Final Energy|Residential and Commercial", "FE|Buildings")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

### Primary Energy

```{r}
items <- c("Primary Energy", "PE")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Oil

```{r}
items <- c("Primary Energy|Oil", "PE|Oil")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Coal

```{r}
items <- c("Primary Energy|Coal", "PE|Coal")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Gas

```{r}
items <- c("Primary Energy|Gas", "PE|Gas")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Nuclear

```{r}
items <- c("Primary Energy|Nuclear", "PE|Nuclear")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Biomass

```{r}
items <- c("Primary Energy|Biomass", "PE|Biomass")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```

#### Primary Energy|Non-Biomass Renewables

```{r}
items <- c("Primary Energy|Non-Biomass Renewables", "PE|Non-Biomass Renewables")
for (s in scenarios) {
  showLinePlots_NGFS(data, items, scen = s)
}
```
